---
title: "Davis_RLCodathon"
author: "BYDavis"
date: "2023-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
species <- read.csv("C:/Users/bydav/Desktop/July2023_SpeciesList.csv", header = TRUE)
library(stringr)
```

# Task Assignments


## Task: way to add species 
Write a little chunk to add new species to the list (and maybe record the row numbers so that downstream scripts and functions can also be run without having to rerun the entire document? Or should it be the other way around - run the to-be-added file first and then append?)

```{r}
# need to add: 
#fundulus heterclitus
#acipenser oxrhynchus

```

## Task: way to search if a species is already present and locate the row if so
## TASK: Try to switch the for loop out for an apply
```{r}

addlist <- c("petromyzon marinus", "anguilla rostrata", "micropterus salmoides", "salvelinus fontinalis", "microgadus tomcod", "salmo salar", "alosa sapidissima", "alosa aestivalis", "morone saxatilis", "morone americana", "alosa pseudoharengus", "osmerus mordax", "sebastes fasciatus", "urophycis tenuis", "brosme brosme", "gadus morhua", "thunnus thynnus", "pollachius virens", "brevoortia tyrannus", "merluccius bilinearis", "oncorhynchus mykiss", "salmo trutta", "acipenser oxrhynchus", "acipenser brevirostrum", "fundulus heterclitus")

for (i in addlist){
  i = str_to_sentence(i)
  print(which(species == i, arr.ind=TRUE))
}

# to find a specific species/entry:
print(which(species == "acipenser oxrhynchus", arr.ind=TRUE))
```

## Task: way to remove a species? - deal with in #5, not now

# Tasks in Review
## Task: Write Zenodo landing page (narrative, links, use cases) - decide on account/hosting

The following files were developed during the Maine eDNA grant (NSF funding info). eDNA (environmental DNA) data include spatial (where the data were collected), environmental (the conditions when the samples were collected), and genetic information, among other categories, but the genetic data is key to understanding and using eDNA to make observations and prediction. When taking an indirect eDNA sample from a river, a mudflat, or beach, it isn't usually obvious where the DNA is coming from and which organisms can be detected. To select which organisms we want to detect, or identify organisms in the sample, we rely on a reference database - a pre-existing, validated collection of sequences that are known to be able to identify a particular species or population. Comparing our results to the reference database or using reference sequences to check if a sample has the target organisms enables eDNA to be used to detect abundance, predict biomass, and in general be used for different scientific questions. 

Unfortunately, the existence and trustworthiness of a reference sequence depends on how much work has been done on an organism in the past, and how much previous scientists have worked to ensure a sequence can be used for that trusted level of identification. For remote areas, cryptic organisms, or organisms which haven't received much genetic attention, a trusted reference sequence may be difficult or impossible to find. This creates difficulties analyzing data, and potentially long delays if a project unexpectedly has to work to find specimens to create proper reference sequences.

The Maine eDNA grant funded numerous projects across nearly every biome and taxonomic group in Maine, and the variable existence or dearth of reference sequences quickly became an apparent issue. The Reference Library Research Group began as an effort to coordinate and consolidate different data workflows so that instead of individuals or separate labs working on different or repeated parts of a problem, the group overall could collaborate and create shared resources that could benefit the entire network of grant members and their collaborators. Over time, the utility of a reference database workflow on other projects and to other groups became more and more apparent, and other examples from initiatives such as CaleDNA gave us ideas on how we might develop tools that could be used for non-Maine eDNA efforts as well.

We developed the following:
(list of outputs in their 'final form')



## Task: Make the credit list
Here, credit categories reference and follow the CRediT system (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7775117/ | http://credit.niso.org/)

Made 2 copies (one for #1-4 and one for ##5) and put them into the Codathon folder on the RefLib drive. Also linked to slack for the codathon people to get first pass at adding their names

## Task: Run full list genera through GGI 
Complete

# Post Codathon Tasks
## Email the RL group about the Zenodo page
## Email the RL group about the codathon - what happened, results, where we are, sending out the Credit pages if people want to join in, and discussion of research forum presentation
## Email the RL species list contributors about the Credit pages and give them second pass at signing up for authorship commitments
## Schedule a research forum presentation time
